Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2012-09-19 chi2012.tex
intro tweaks and foo bar

In recent years, an unprecedented quantity and variety of information has been made available as structured data on the Web, through APIs, datasets, and data feeds.  This information includes: access to data previously hidden behind services and applications, such as retailer product catalogues; information not previously directly released to the public, such as open government data or records of personal financial transactions; and new kinds of data generated by emerging kinds of data sources, such as wearable and smartphone-based biosensors.  A primary goal of opening up this data is that end-user citizens can make more informed decisions pertaining to their health, wealth, and well-being \cite{}.  This data is predominately used by specialised app developers, journalists and other ``data specialists'' but remains inaccessible to citizens because of the available tools.  For example, the most commonly used structured personal information management (PIM) tools either manage a small, fixed set of data types (such as digital calendaring tools and  to-do list managers) or provide little or no support for structured data at all, such as text editors and sketching/drawing tools \cite{}.

@@

For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, for music sites this was song info, etc.  Records were located either by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query if available.  For each obtained data record, two simple complexity metrics were computed: an \emph{average width} and \emph{average depth} of record.  For determining the width, we first examined the record to find the appropriate level for comparison; for example, Amazon's retail product ``Item'' root record had few attributes about the actual product compared to its ``ItemAttributes'' sub-property, which was used instead. Thus, all comparisons were made to maximize comparative similarity.  Depth was computed as the average number of levels from record root to leaves across all properties, and was measured, with width, to convey the overall size/complexity of each record.

@@

Once the equivalent properties were established, we look at how many of these properties were terminologically inconsistent, that is, identified using different property names.   Then, ignoring property name, we looked to see if its values were structurally comparable.  If two fields were of the same data type and, if complex, had equivalent subfields, they were considered comparable; otherwise they were considered \emph{structurally inconsistent}.  An example of such a structural inconsistency were``birthday'' fields in Facebook and LinkedIn; Facebook represented it as a single string formatted ``MM/DD/YYYY'', while LinkedIn represented it as a sub-structure with separate ``Month'', ``Day'' and ``Year'' subfields.

Finally, we looked at fields that were equivalent, yet which could not be compared for reasons other than the terminological and structural reasons above.  These \emph{modelling differences} were caused by a variety of reasons including, but not limited to, the following: measurement unit inconsistencies, how data was measured (e.g., dewpoint vs relative humidity), differences in scale or granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').

@@

Looking at equivalent properties, we foun overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.  While some of these 'unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply not present.  Only one of the music data providers, Spotify, included information about the album(s) that featured a particular song.  Such omissions show us that these data APIs reflect each service provider's own capabilities and internal data representations very specifically, and that integrating data such records from multiple sources could allow more complete and useful representations to be assembled.

The second finding is that equivalent properties were sometimes, though only rarely, given the same name.  Thus, terminological heterogeneity was prevalent across data records we observed.  Structural heterogeneity, however, was comparatively rare, with the most common case being similar to the birthday example above; a leaf property on one record modeled as a subtree in the other.  We also encountered only 2 examples of fields which had data nested in textual description fields, both from records in the Weather category, which hints their intended use in RSS and other feed readers than as stuctured data feeds.

Examples of modelling/semantic heterogeneity instances were rarer than those of structural heterogeneity, with the few we found mostly being attributed to concepts on one site not mapping perfectly to those of another site; ``Promotions'' versus ``Offers'',

Although not included in the table above, we also examined the ranges of enumerated type values across data sources.  There was substantial terminological heterogeneity among such values, but also many cases where one source had additional values not supported in the range of the other.  For example, the property ``Gender''  on Facebook allowed only ``Male'' and ``Female'' for values, while Google+ included a third option,``other''.  ``Relationship status'' was another which was similarly complex. Amazon's

>>>>

In recent years, an unprecedented quantity and variety of information has been made available as structured data on the Web.  This information has come by way of many different channels, from the various APIs now offered by most of the major online retailers, apps, sharing and service platforms, the massive data sets published by governmental and private institutions, to new streams being produced by various kinds of emerging data sources, such as wearable and smartphone-based biosensors.  A primary goal of making much of this data available has been to give end-user citizens the ability to make more informed decisions pertaining to their health, wealth, and well-being \cite{Shadbolt:2006:SWR:1155313.1155373}.  This data is predominately used by specialised app developers, journalists and other ``data specialists'' but remains inaccessible to citizens because of the available tools.  For example, the most commonly used structured personal information management (PIM) tools either manage a small, fixed set of data types (such as digital calendaring tools and  to-do list managers) or provide little or no support for structured data at all, such as text editors and sketching/drawing tools \cite{}.

@@

For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, for music sites this was song info, etc.  Records were located either by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query if available.  For each obtained data record, two simple complexity metrics were computed: an \emph{average width} and \emph{average depth} of record.  For determining the width, we first examined the record to find the appropriate level for comparison; for example, Amazon's retail product ``Item'' root record had few attributes about the actual product compared to its ``ItemAttributes'' sub-property, which was used instead. Thus, all comparisons were made to maximize comparative similarity.  
% Depth was computed as the average number of levels from record root to leaves across all properties, and was measured, with width, to convey the overall size/complexity of each record.

@@

Once the equivalent properties were established, we counted the properties that were terminologically inconsistent, that is, named differently between records.   Then, ignoring property name, we looked to see if its values were structurally comparable.  If two fields were of the same data type and, if complex, had equivalent subfields, they were considered comparable; otherwise they were considered \emph{structurally inconsistent}.  An example of such a structural inconsistency were``birthday'' fields in Facebook and LinkedIn; Facebook represented it as a single string formatted ``MM/DD/YYYY'', while LinkedIn represented it as a sub-structure with separate ``Month'', ``Day'' and ``Year'' subfields.

Finally, we looked at fields that were equivalent, yet could not be compared for reasons other than the two reasons above.  These \emph{modelling differences} were had a variety of causes, including, but not limited to, the following: measurement unit inconsistencies, measurement method (e.g., dewpoint vs relative humidity), differences in scale and granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').

@@

Looking at equivalent properties, we foun overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.  While some of these 'unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply absent.  Only one Spotify, among music providers, included information about the album(s) featuring a given song.  Such omissions show us that these data APIs reflect each provider's own specific capabilities and needs, and suggests that integrating information from multiple sources may allow more complete and useful representations to be assembled.

The second finding is that equivalent properties were only rarely given the same name.  Thus, terminological heterogeneity was prevalent across data records we observed.  Structural heterogeneity, however, was comparatively rare, with the most common case being similar to the birthday example above; a leaf property on one record modeled as a subtree in the other.  We also encountered only 2 examples where data was nested in textual description fields, both from records in the Weather category, which hints their intended use in RSS feed readers rather than in stuctured database applications.  Examples of modelling/semantic heterogeneity instances were rarer than those of structural heterogeneity, with the few we found mostly being attributed to concepts on one site not mapping perfectly to those of another site, and mismatches in units of measurement (commonly found in the Weather forecast measurements).

Although not included in the table above, we also examined the source schemas of many of hte data sources for property enumerated type values.  There was substantial terminological inconsistency among such values, but also cases where some sources had values not represented in the other source's sets.  For example, the property ``Gender''  on Facebook allowed only supported ``Male'' and ``Female'' for values, while Google+ included a third option,``other''.  ``Relationship status'' was another which was similarly complex. None of the string representations of enumerated value sets matched across data sources perfectly.

In summary, while data records ranged in complexity, no data source subsumed the rest; all contained properties that other sources did not have.  Moreover, the most common types of heterogeneity exhibited were terminological and value-related, although simpler kinds of structural hetereogeneity (pertaining to representing substructures of values) were also common.
